White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-10 papers/wdmerger_II/paper.tex
Fix the resolution number
Clarify wording regarding iso7

We have added to the \wdmerger\ problem setup the ability to take advantage of 
the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
in this study, because the axisymmetry inherent to the cylindrical coordinate system
lends itself well to the head-on collision problem. In this coordinate system, we
align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
in both dimensions, so that there are twice as many zones in the $z$ dimension in
a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a resolution
of 156.25 km.

@@

\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
the region at the central contact point does not fully stop burning, yielding a quasistatic energy
release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
hundred it normally takes). The full effect of these additional timesteps only amounts to a trace change
in the nickel production. By examining the behavior of zones near the collision point, we discovered the
reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.

>>>>

We have added to the \wdmerger\ problem setup the ability to take advantage of 
the 2D cylindrical ($R-z$) coordinate system evolution in \castro, and we use it
in this study, because the axisymmetry inherent to the cylindrical coordinate system
lends itself well to the head-on collision problem. In this coordinate system, we
align the WDs along the $z$-axis (which is analogous to the $x$-axis in Cartesian
evolution), with the center of the WDs at $R = 0$. The un-simulated $\phi$ dimension
then would extend the WDs through a $2\pi$ revolution. The domain has width $4 \times 10^{9}$
cm along the $z$ axis, and is half as wide along the $R$ axis. The resolution is equal
in both dimensions, so that there are twice as many zones in the $z$ dimension in
a uniform grid. The default grid has 256 zones along the $z$ axis, yielding a
spatial resolution of 312.5 km.

@@

\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
the region at the central contact point does not fully stop burning, yielding a quasistatic energy
release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
hundred it normally takes, using our designated stopping criterion). The full effect of these additional
timesteps only amounts to a trace change in the nickel production. By examining the behavior of zones
near the collision point, we discovered the reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.
